A Philosophy of (Inductive) Testing

### 2024-03-08 mss.tex
Post semifinal draft.

\begin{abstract}
Testing, including unit testing, forms a critical step in software development.  This article explores the philosophy and practice of testing in the context of Hoon and Urbit development, particularly as motivated by the development of urQL and Obelisk.  Inductive testing provides an efficient yet complete basis for verifying code behavior.  Principles and best practices are suggested, as well as a practical example of testing in Hoon.
\end{abstract}

@@

We were intentionally dismissive of testing methodologies above because they are mere conventions. There is however some actual theory we can leverage in figuring out what and how to test.\footnote{\citet{WikipediaInduction} provides a modestly technical exposition of mathematical induction. You can skip it because we will explain it non-rigorously for lay programmers. We also quibble about the inclusion of zero as a natural number. While old school maths started the natural numbers with “1”, but computer science has since infected maths. Have you ever seen zero of anything? No—it is not natural at all. Zero is a very, very special number. It's an abstraction that does not map to anything in the physical world.}

Peano arithmetic is a better lay programmer's introduction to proof by induction. There exists (or perhaps \emph{does not exist}) a special concept, zero. Playing fast and loose with classical logic, “ex nihil sequitur quodlibet”, from nothing (more commonly a falsehood, or a contradiction) follows everything (i.e. anything).\footnote{The “principle of explosion” rigorously follows from use of disjunctive syllogisms.  Here, we jocosely indicate that by proceeding from zero inductively we can demonstrate desired properties of testing.} Zero is not only an integer; rather, in terms of Nock-based programming, \lstinline[style=inlinecode]{~} is not only the beginning of counting,\footnote{Yarvin opted to invert \texttt{true} (as 0) and \texttt{false} (as 1) \citep{Yarvin2006}, meaning that Nock loobeans do not align inductively with other inductive types.} but the nothing of every inductive type, most importantly trees and lists. When speaking to programmers, Peano tells us there is a universal nothing \lstinline[style=inlinecode]{~} and there is a function called successor $\textrm{Succ}$, which produces some next thing from a previous thing (or lack of thing). Applying $\textrm{Succ}$ to nothing, $\textrm{Succ}(0)$, gives us the first thing.  Applying $\textrm{Succ}$ to the first thing, $\textrm{Succ}(\textrm{Succ}(0))$ gives us the second thing, and so on.

@@

From our experience with the F\# implementation of property-based testing, we expect that you will need to boost the number of random input permutations beyond the default of 100 to get the kind of coverage that is ``reasonably'' exhaustive. We found 10,000 to be frequently adequate. (Since the inputs are randomly generated each run, however, a property test of production code may fail when nothing has changed. Then think about how to explain to your boss that your tests are not deterministic.)

@@

\printbibliography
\end{document}

>>>>

\begin{abstract}
  \noindent
  Testing, including unit testing, forms a critical step in software development.  This article explores the philosophy and practice of testing in the context of Hoon and Urbit development, particularly as motivated by the development of urQL and Obelisk.  Inductive testing provides an efficient yet complete basis for verifying code behavior.  Principles and best practices are suggested, as well as a practical example of testing in Hoon.
\end{abstract}

@@

We were intentionally dismissive of testing methodologies a\-bove because they are mere conventions. There is however some actual theory we can leverage in figuring out what and how to test.\footnote{\citet{WikipediaInduction} provides a modestly technical exposition of mathematical induction. You can skip it because we will explain it non-rigorously for lay programmers. We also quibble about the inclusion of zero as a natural number. While old school maths started the natural numbers with “1”, but computer science has since infected maths. Have you ever seen zero of anything? No—it is not natural at all. Zero is a very, very special number. It's an abstraction that does not map to anything in the physical world.}

Peano arithmetic is a better lay programmer's introduction to proof by induction. There exists (or perhaps \emph{does not exist}) a special concept, zero. Playing fast and loose with classical logic, “ex nihil sequitur quodlibet”, from nothing (more commonly a falsehood, or a contradiction) follows everything (i.e. anything).\footnote{The “principle of explosion” rigorously follows from use of disjunctive syllogisms.  Here, we jocosely indicate that by proceeding from zero inductively we can demonstrate desired properties of testing.} Zero is not only an integer; rather, in terms of Nock-based programming, \lstinline[style=inlinecode]{~} is not only the beginning of counting,\footnote{Yarvin opted to invert \texttt{true} (as 0) and \texttt{false} (as 1) \citep{Yarvin2006}, meaning that Nock loobeans do not align inductively with other inductive types.} but the nothing of every inductive type, most importantly trees and lists. In speaking to programmers, Peano tells us there is a universal nothing \lstinline[style=inlinecode]{~} and there is a function called successor $\textrm{Succ}$, which produces some next thing from a previous thing (or lack of thing). Applying $\textrm{Succ}$ to nothing, $\textrm{Succ}(0)$, gives us the first thing.  Applying $\textrm{Succ}$ to the first thing, $\textrm{Succ}(\textrm{Succ}(0))$ gives us the second thing, and so on.

@@

From our experience with the F{$^\sharp$} implementation of pro\-perty-based testing, we expect that you will need to boost the number of random input permutations beyond the default of 100 to get the kind of coverage that is ``reasonably'' exhaustive. We found 10,000 to be frequently adequate. (Since the inputs are randomly generated each run, however, a property test of production code may fail when nothing has changed. Then think about how to explain to your boss that your tests are not deterministic.)

@@

\selectlanguage{USenglish}
\printbibliography
\end{document}
